---
title: "Additional Analyses"
output: 
  rmarkdown::html_vignette:
    toc: yes
    number_sections: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{7. Additional Analyses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', fig.height = 4, 
                      fig.width = 5)
```
```{r, echo = FALSE}
e <- TRUE
e2 <- FALSE
```


# Overview

Additional analyses undertaken to evaluate changes in scale by year, and to compare scales estimated in this study with those from several earlier studies. 

# Analyses
## Data and libraries

```{r, message = FALSE, warning=FALSE}
library(ecoscales)
library(viridis)
library(ggplot2)
data("dimbreaks")
data("bootperturb")
data("datf")
```

## Observational methods

### Percent by observational category
```{r}
knitr::kable(datf[, list("%" = round(.N / nrow(datf) * 100, 2)), 
                    by = study_type])
```

<a href="#top">Back to top</a>

### Remote/automated/paleo shares of observations

Comparing the share of these observational types relative to certain extent and durational thresholds
```{r, eval = e}
# share of observations made with remote/other above 10,000 ha extent
aext <- 10^4
a <- datf[, {
  a = length(which(act_ext > aext))
  b1 = length(which((act_ext > aext) & like(study_type, "remote")))
  b2 = length(which((act_ext > aext) & like(study_type, "other")))
  round(c("remote" = b1 / a * 100, "other" = b2 / a * 100), 1)
}]
knitr::kable(t(as.matrix(a)))

# share of automated/paleo observations of observatons lasting more than 1 year
adur <- 365
b <- datf[, {
  a = length(which(act_dur > adur))
  b1 = length(which((act_dur > adur) & like(study_type, "auto")))
  b2 = length(which((act_dur > adur) & like(study_type, "paleo")))
  round(c("automated" = b1 / a * 100, "paleo" = b2 / a * 100), 1)
}]
knitr::kable(t(as.matrix(b)))

# share of automated/paleo observations of observatons lasting < 1 year
d <- datf[, {
  a = length(which(act_dur <= adur))
  b1 = length(which((act_dur <= adur) & like(study_type, "auto")))
  b2 = length(which((act_dur <= adur) & like(study_type, "paleo")))
  round(c("automated" = b1 / a * 100, "paleo" = b2 / a * 100), 1)
}]
knitr::kable(t(as.matrix(d)))

```

<a href="#top">Back to top</a>

## Annual changes in observational scales

### Study type by year

#### Calculation

Calculate types of study by year, as total and percent
```{r, warning = FALSE, message = FALSE, eval = e}
# One Oecologia study listed as 2004 (by Endnote) was actually 2003 
# Set it to 2004
stypes <- copy(datf)  # new dataset out of dat (which retain year info)
stypes[study_year < 2004, study_year := 2004] 

# ID NA eff_ext values
na_effext <- datf[is.na(eff_ext), unique(DOI)]

# Estimate and plot study type by year
stypeyr <- stypes[, list(.N, study_type), by = study_year][, {
  list("tot" = N, "ct" = .N, "prop" = .N / N * 100)
}, by = .(study_type, study_year)][order(study_year)]
stypeyr <- stypeyr[, lapply(.SD, mean), .SDcols = c("tot", "ct", "prop"), 
                   by = .(study_type, study_year)]
```

#### Figure 
```{r, eval = e2}
obs_types <- c("remote", "field", "auto")
obs_types2 <- c("Remote sensing", "Field", "Automatic")
cols <- c("orange", "green4", "blue4")
png("figures/stype_by_yr.png", width = 7, height = 2.75, res = 300, 
    units = "in")
par(mfrow = c(1, 3), oma = c(0, 3, 0, 0), mar = c(4, 1, 2, 1))
for(i in 1:3) {  # i <- 1
  plot(c(2004, 2014), c(0, 100), xlab = "", ylab = "", pch = "")
  stypeyr[like(study_type, obs_types[i]), {
    points(study_year, prop, col = cols[i], pch = 20)
    yr = as.numeric(study_year)
    slm = lm(prop ~ yr, weights = tot)  # weighted regression
    slms = summary(slm)
    abline(slm, col = cols[i])
    cf = round(coef(slm)[1:2], 4)
    pf = round(coef(slms)[8], 2)
    rf = round(slms$adj.r.squared, 2)
    cfa = paste0(cf[2], "% year")
    text(x = 2006, y = 100, labels = substitute(x^-1, list(x = cfa)), 
         cex = 0.9)
    text(x = 2011.5, y = 100, cex = 0.9,
         labels = substitute(R^2~"="~x~"; p <"~y, list(x = rf, y = pf)))
    if(i == 1) mtext("percent", side = 2, line = 2.5, cex = 0.8)
    if(i == 2) mtext("year", side = 1, line = 2.5, cex = 0.8)
    mtext(obs_types2[i], side = 3, cex = 0.8)
  }]
}
o <- dev.off()
```

#### Extrapolation to 2017

Extrapolating percentage of studies using remote sensing to 2017
```{r, eval = e}
preds <- stypeyr[like(study_type, "remote"), {
  slm = lm(prop ~ study_year, weights = tot)
  predict(slm, newdata = data.frame("study_year" = 2004:2017))
}]
names(preds) <- 2004:2017
```

Percent of remote sensing studies by year
```{r, eval = e}
knitr::kable(t(as.matrix(round(preds, 1))))
```


Average annual percent of studies using remote sensing 2004-2014, and projected increases in this percent through 2017 (2004-2017) 
```{r, eval = e}
# murs2014 <- stypeyr[study_type == "remote", round(mean(prop), 2)]
murs2014 <- stypeyr[, sum(ct[study_type == "remote"]) / sum(ct) * 100]
pctinc <- mean(preds) / mean(preds[1:(length(preds) - 3)])

pcttab <- cbind("2014" = murs2014, "pct increase" = (pctinc - 1) * 100,
                "2017" = murs2014 * pctinc)
knitr::kable(round(pcttab, 1))
```

```{r, eval = e}
types <- datf[, .N, by = study_type]
rs2017 <- mean(preds)  # avg predicted % using RS through 2004-2017
rs2014 <- mean(preds[1:(length(preds) - 3)])  # avg predicted % RS 2004-2014
# rs2017 / rs2014

# translate to N more RS studies would have been selected given those %ages
rsinc <- rs2017 / rs2014 * types[study_type == "remote", N] - 
                 types[study_type == "remote", N]  

remote <- datf[study_type == "remote", eff_ext]
notremote <- datf[study_type != "remote", eff_ext]

set.seed(1)
ext2017 <- unlist(lapply(1:1000, function(x) {  # x = 1
  ind <- sample(1:length(remote), size = round(rsinc))
  mean(c(remote, remote[ind], notremote), na.rm = TRUE)
}))

dec1417 <- log10(mean(ext2017, na.rm = TRUE) / 
                   datf[, mean(eff_ext, na.rm = TRUE)])
qtiles1417 <- log10(quantile(ext2017, c(0.025, 0.975), na.rm = TRUE) / 
                      datf[, mean(eff_ext, na.rm = TRUE)])

pctdiff1417 <- mean(ext2017, na.rm = TRUE) / datf[, mean(eff_ext, na.rm = TRUE)]
pctqtiles1417 <- quantile(ext2017, c(0.025, 0.975), na.rm = TRUE) / 
                      datf[, mean(eff_ext, na.rm = TRUE)]

otab <- cbind("decade" = round(dec1417, 4), "dec2.5" = round(qtiles1417[1], 4),
              "dec97.5" = round(qtiles1417[2], 4),
              "pctdiff" = round((pctdiff1417 - 1) * 100, 1),
              "pctdiff2.5" = round((pctqtiles1417[1] - 1) * 100, 1), 
              "pctdiff97.5" = round((pctqtiles1417[2] - 1) * 100, 1))
knitr::kable(otab)
```

<a href="#top">Back to top</a>

### All scales

#### Figure
```{r, eval = e2}
vars <- c("plot_res", "eff_ext", "t_btwn_samp", "eff_dur")
varlabs <- c("Resolution", "Extent", "Interval", "Duration")
labcols <- c(brewer.pal(8, "Reds")[c(7, 7)], brewer.pal(8, "Blues")[c(7, 7)])
png("figures/dim_by_yr.png", width = 7, height = 6, res = 300, 
    units = "in")
par(mfrow = c(2, 2), mar = c(1, 3, 1, 1), oma = c(2, 3, 0, 0))
for(i in 1:length(varlabs)) {  # i <- 1
  if(i == 3) {
    DT <- copy(stypes[t_btwn_samp != 0, ])
  } else {
    DT <- copy(stypes)
  }
  DT[, list(mean(get(vars[i]), na.rm = TRUE), .N), by = study_year][, {
    dim = log10(V1)
    yr = as.numeric(study_year)
    if(i %in% 1:2) rng = c(0, 10)
    if(i %in% 3:4) rng = c(0, 5)
    if(i %in% 3:4) {
      xlabs = seq(2004, 2014, 2)
    } else {
      xlabs = rep("", 6)
    }
    plot(yr, dim, ylim = rng, xaxt = "n", pch = 20, col = labcols[i], 
         main = varlabs[i], mgp = c(2, 0.5, 0), tcl = -0.2, ylab = "")
    axis(1, at = seq(2004, 2014, 2), labels = xlabs, tcl = -0.2, 
         mgp = c(2, 0.5, 0))
    slm = lm(dim ~ yr, weights = N)
    slms = summary(slm)
    abline(slm, col = labcols[i])
    cf = round(coef(slm)[1:2], 3)
    pf = round(coef(slms)[8], 2)
    rf = round(slms$adj.r.squared, 2)
    text(x = 2006.25, y = rng[2] * 0.97, cex = 0.9,
         labels = substitute(x~year^-1, list(x = cf[2])))
    text(x = 2011.5, y = rng[2] * 0.97, cex = 0.9,
         labels = substitute(R^2~"="~x~"; p <"~y, list(x = rf, y = pf)))
    if(i == 1) mtext(expression("log10 m"^2), side = 2, line = 2)
    if(i == 2) mtext("log10 ha", side = 2, line = 2)
    if(i %in% 3:4) mtext("log10 days", side = 2, line = 2)
  }]
}
o <- dev.off()
```

#### Projected increase in mean observational extent

Between 2014-2017
```{r, eval = e}
preds <- stypes[, list(mean(eff_ext, na.rm = TRUE), .N), by = study_year][, {
  dim = log10(V1)
  yr = as.numeric(study_year)
  slm = lm(dim ~ yr, weights = N)
  predict(slm, newdata = data.frame("yr" = 2004:2017))
}]
# preds[14] - preds[13]

# projected percent increase in extent 2014-2017
pctinc <- mean(preds) / mean(preds[1:(length(preds) - 3)])
muext2014 <- stypes[, mean(eff_ext, na.rm = TRUE)]
pcttab <- cbind("2014" = log10(muext2014), "pct increase" = (pctinc - 1) * 100,
                "2017" = log10(muext2014 * pctinc), 
                "decade" = log10((pctinc * muext2014) / muext2014))
knitr::kable(round(pcttab, 3))
```

<a href="#top">Back to top</a>

## Scales in earlier studies
### Duration from Tilman, 1989
```{r, eval = e}
kdat <- rbindlist(bootperturb)
std <- function(x) (x - min(x)) / diff(range(x))

# Tilman (1989), Figure 6.1 (visually interpreted)
# Data extracted from figure using http://arohatgi.info/WebPlotDigitizer/app/
f <- system.file("extdata/tilman1989.csv", package = "ecoscales")
tilman <- fread(f)

## Find n and proportion of non-experimental observations per year
# first find mean/median value of data > 100 years in our observed data
gt100 <- kdat[eff_dur > (365 * 100), {
  list("med" = median(eff_dur) / 365, "mu" = mean(eff_dur) / 365) 
}]

# find weights and proportion of studies lasting particular duration in Tilman
tilman_noexp <- tilman[, {
  a = round(623 * prop[fig == 1] * (1 / sum(prop[fig == 1])))
  b = round(180 * prop[fig == 2] * (1 / sum(prop[fig == 2])))
  padj = a - b
  w = padj / sum(padj)
  list("yrmu" = yrmu[fig == 1], "n" = padj, "prop" = w)
}]

# median and mean duration in Tilman data, assuming 150 years for studies > 100
tilmedmu <- tilman_noexp[, rep(yrmu, n), by = n][, {
  list("med" = median(V1), "mean" = mean(V1))
}]

# as above, but testing different durations for studies lasting > 100 years
gt1002 <- unname(c(200, 500, 1000, unlist(gt100)))
tilmedmu2 <- rbindlist(lapply(gt1002, function(x) {
  DT2 = copy(tilman_noexp)[yrmu == 150, yrmu := x]
  DT2[, rep(yrmu, n), by = n][, list("med" = median(V1), "mean" = mean(V1))]
}))

# And calculate only for years less than 100
tilmedmu_lt100 <- tilman_noexp[yrmu < 150, rep(yrmu, n), by = n][, {
  list("med" = median(V1), "mean" = mean(V1))
}]

# calculation mean and median duration from our data, both from replicated 
# and non-replicated studies
# including studies > 100 years
durmedmu <- kdat[, {
  list("med" = median(eff_dur[t_btwn_samp != (365 * 10000)], na.rm=TRUE) / 365, 
       "mean" = mean(eff_dur[t_btwn_samp != (365 * 10000)], na.rm = TRUE) / 365,
       "medall" = median(eff_dur, na.rm = TRUE) / 365, 
       "meanall" = mean(eff_dur, na.rm = TRUE) / 365)
}]

# excluding studies > 100 years
durmedmu_lt100 <- kdat[eff_dur < (365 * 100), {
  list("med" = median(eff_dur[t_btwn_samp != (365 * 10000)], na.rm=TRUE) / 365, 
       "mean" = mean(eff_dur[t_btwn_samp != (365 * 10000)], na.rm = TRUE) / 365,
       "medall" = median(eff_dur, na.rm = TRUE) / 365, 
       "meanall" = mean(eff_dur, na.rm = TRUE) / 365)
}]
```

Median, mean duration from Tilman (1989), varying according to assumption of duration of studies > 100 years
```{r, eval = e}
yr <- c(150, 200, 500, 1000, unlist(gt100))
knitr::kable(cbind("max_years" = round(yr), 
                   round(rbind(tilmedmu, tilmedmu2), 2)))
```

Tilman with studies > 100 years dropped
```{r, eval = e}
knitr::kable(round(tilmedmu_lt100, 2))
```

Tilman (1989) duration of observational studies by bin
```{r, eval = e}
knitr::kable(tilman_noexp[, {
  a = round(prop * 100, 1)
  names(a) = tilman[fig == 1, yr]
  rbind(a, n)
}])
```

Median, mean duration from this study
```{r, eval = e}
knitr::kable(round(durmedmu, 2))
```

This study with years > 100 dropped
```{r, eval = e}
knitr::kable(round(durmedmu_lt100, 2))
```

<a href="#top">Back to top</a>

### Kareiva and Anderson (1989)

Kareiva and Anderson (1989) report resolution data. The 80th, 90th, and 95th percentile values of resolutions in Kareiva and Anderson are also calculated as a basis for comparison.
```{r, eval = e}
# helper func
q80 <- function(x) quantile(x, 0.8, na.rm = TRUE)
q90 <- function(x) quantile(x, 0.9, na.rm = TRUE)
q95 <- function(x) quantile(x, 0.95, na.rm = TRUE)

f <- system.file("extdata/kareiva_anderson1989.csv", package = "ecoscales")
kareiva <- fread(f)
a <- kareiva[, {
  list("med" = median(res^2), "mean" = mean(res^2), "80" = q80(res^2), 
       "90" = q90(res^2), "95" = q95(res^2))
}]
b <- kdat[, {
  list("med" = median(plot_res), "mean" = mean(plot_res), "80" = q80(plot_res), 
       "90" = q90(plot_res), "95" = q95(plot_res))
}]
study <- c("Kareiva & Anderson, 1989", "This study")
knitr::kable(cbind(study, round(rbind(a, b), 2)))
```

<a href="#top">Back to top</a>

### Porter et al (2005)

Porter et al (2005) report interval and extent data extracted from a number of studies.

#### Interval
```{r, eval = e}
f <- system.file("extdata/porter.csv", package = "ecoscales")
porter <- fread(f)
porter <- porter[sample == 1]
porter[, ext := ext^2 / 10000]

study <- c("Porter et al, 1989", "This study")
a <- porter[int > 0, {
  list("med" = median(int), "mean" = mean(int), "80" = q80(int), "90" = q90(int),
       "95" = q95(int))
}]

b <- kdat[t_btwn_samp != (365 * 10000), {
  list("med" = median(t_btwn_samp), "mean" = mean(t_btwn_samp), 
       "80" = q80(t_btwn_samp), "90" = q90(t_btwn_samp), "95" = q95(t_btwn_samp))
}]
knitr::kable(cbind(study, round(rbind(a, b))))
```

#### Extent
```{r, eval = e}
study <- c("Porter et al, 1989", "This study", "Ratio")
a <- porter[, {
  list("med" = median(ext), "mean" = mean(ext), "80" = q80(ext), 
       "90" = q90(ext), "95" = q95(ext))
}] 
b <- kdat[, {
  list("med" = median(eff_ext, na.rm = TRUE), 
       "mean" = mean(eff_ext, na.rm = TRUE), "80" = q80(eff_ext), 
       "90" = q90(eff_ext), "95" = q95(eff_ext))
}]
knitr::kable(cbind(study, round(rbind(a, b, b / a))))
```

<a href="#top">Back to top</a>


