---
title: "Main Analyses - Figure 3"
output: 
  rmarkdown::html_vignette:
    toc: yes
    number_sections: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{6. Main Analyses - Figure 3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', fig.height = 4, 
                      fig.width = 6, cache = TRUE)
```
```{r, echo=FALSE}
e <- TRUE
e2 <- TRUE
```

# Overview

The resampled full dataset created in the [Data Pre-processing vignette](data-pre-process.html) is used to calculate and plot the magnitude of difference extent and actual extent and duration and actual duration.

# Analyses
## Data and libraries

```{r, message = FALSE, warning=FALSE}
library(ecoscales)
library(viridis)
library(ggplot2)
data("dimbreaks")
data("bootperturb")
# data(datf)
```

## Actual versus effective duration/extent

Prepare the data for creating boxplots: calculate differences between log-transformed actual and effective extents/durations, assign the values to different bins (intervals) of extent/duration, calculate the percentage of observations in each bin. 
```{r, eval = e}
bootp_dt <- rbindlist(bootperturb)
kdat <- copy(bootp_dt)  # bootstrap
# kdat <- copy(datf)

# limits
xyvars <- list(c(x = "eff_ext", y = "act_ext"),  # variable pairs 
               c(x = "eff_dur", y = "act_dur"))
lims <- list("eff_ext" = range(aaxis2$logarea), # limits
             "act_ext" = range(aaxis2$logarea), 
             "eff_dur" = range(taxis2$logdays[-5]), 
             "act_dur" = range(taxis2$logdays[-5]))
ivals <- list("ext" = aaxis2$logarea, "dur" = taxis2$logdays[-5])
dropones <- c(FALSE, TRUE)
onms <- list(c("eext", "aext"), c("edur", "adur"))
wgtint <- c(0, 1, seq(5, 40, 5)) / 100
wgttxt <- (c("  0", "  1", "  5", seq(10, 40, 5)))

# types
types <- c("field|remote|paleo|automated|other", "field", "automated", 
           "remote|other", "paleo")
pctints <- c(0, 1, seq(5, 100, 5))

# calculate difference magnitudes by type
dif_ints <- lapply(1:length(types), function(i) {  # i <- 3
  # print(i)
  dif_int <- lapply(1:length(xyvars), function(x) {  # x <- 1
    # print(x)
    xyv <- xyvars[[x]]
    onm <- onms[[x]]
    ival <- ivals[[x]]
    xl <- lims[[xyv["x"]]]
    yl <- lims[[xyv["y"]]]
    
    # grab all data first (want to preserve total N obs), remove unrepl as
    # needed and subset down to variables, dropping NAs
    dat <- copy(kdat)
    # setnames(dat, "study_type", "type")
    if(x == 2) dat <- dat[t_btwn_samp != 365 * 10000, ]  # drop unreplicated
    dat <- dat[, list("type" = type, "x" = get(xyv["x"]), "y" = get(xyv["y"]))]
    dat <- na.omit(dat)

    # subset by type
    dfd <- dat[like(type, types[i])]
    dfd[type == "other", type := "remote"]
    type <- dfd$type
    
    # type[type == "other"] <- "remote"
    dfd <- dfd[, lapply(.SD, log10), .SDcol = c("x", "y")]
    dfd <- cbind(type, dfd)
    
    # dfd <- cbind(type, dfd)
    dfd[x < xl[1], x := xl[1]]  # confine x to min
    dfd[x > xl[2], x := xl[2]]  # confine x to max
    dfd[y < yl[1], y := yl[1]]  # confine y to min
    dfd[y > yl[2], y := yl[2]]  # confine y to max
    # dfd <- na.omit(dfd)
    
    # difference and intervals - if want to plot versus actual dur/ext, switch
    # x for y in findInterval
    dfd[, dif := x - y]
    dfd[, int := findInterval(x, ival, rightmost.closed = TRUE)]
    setnames(dfd, c("x", "y"), onm)
    
    # plot related variables, finding percentages in each bin, 
    # occupied intervals 
    ints <- sort(dfd[, unique(int)])
    nobs <- sapply(ints, function(i) dfd[int == i, .N])
    # wgt <- nobs / sum(nobs)
    wgt <- nobs / nrow(dat) * 100
    intmax <- findInterval(max(wgt), pctints) + 1  # upper int
    wgti <- pctints[1:intmax]  # select all intervals
    # wgti <- c(0, 1, seq(5, max(round(wgt, 1)) * 100, 5))
    cols <- findInterval(wgt, wgti)  # corresponding color index
    return(list("db" = dfd, "wgt" = wgt, "wgti" = wgti,
                "col" = cols, "xl" = xl, "yl" = yl, "ival" = ival, 
                "ints" = ints))
  })
  names(dif_int) <- c("e", "d")
  dif_int
})
names(dif_ints) <- c("all", "field", "automated", "remote", "paleo")
```

## Figure 3

Box plots by bins of extent/duration, colored and labelled by the percent of observations per bin.

```{r, eval = e2}
# y grid line function
ylines <- function(yr, by, lty = 1, col = "grey") {
  abline(h = seq.int(from = yr[1], to = yr[2], by = 2), lty = lty, col = col)
}

cxa <- 0.7
ayl <- c(0, 20)
byl <- c(0, 40)

colpals <- c("Blues", "Reds")
pal_len <- c(5, 9)  # max ncolors in palette
ylabs <- c("Extent", "Duration")
a2mgp <- c(3, 0.5, 0)
bwex <- c(0.9, 1.4)


png("figures/fig3.png", width = 7, height = 4, res = 600, 
    units = "in")
par(mfrow = c(1, 2), mar = c(4, 0, 1, 0.75), oma = c(0, 3, 0, 0))
for(i in 1:2) {  # i <- 1
  di <- dif_ints$all[[i]]  # subset list element
  bxa <- rowMeans(cbind(di$ival[-length(di$ival)], di$ival[-1]))[di$ints]
  bxw <- (di$ival[-1] - di$ival[-length(di$ival)])[di$ints]
  cols <- brewer.pal(pal_len[i], colpals[[i]])
  mus <- di$db[order(int), list("mu" = mean(dif)), by = int]

  yls <- c(0, 14)
  plot(di$xl, yls, pch = "", xlab = "", ylab = "", xaxt = "n", yaxt = "n",
       mgp = c(3, 0.5, 0), tcl = -0.2, xaxs = "i")
  polygon(c(di$xl, rev(di$xl), di$xl[1]), c(-1, -1, rep(yls[2] * 1.2, 2), -1), 
          col = "grey90")
  ylines(yr = yls, by = 2, col = "grey95")
  b <- boxplot(formula = dif ~ int, data = di$db, col = cols[di$col], 
               range = 0, width = bxw, at = bxa, xlim = di$xl, boxwex = bwex[i], 
               axes = FALSE, outline = FALSE, lty = 1, add = TRUE, 
               staplewex = 0.2)
  points(bxa, mus$mu, pch = 20, col = "black")
  text(bxa, b$stats[5, ] + 0.15, round(di$wgt, 1), cex = 0.3)
  
  axis(1, at = ivals[[i]], labels = list(alab2, tlab2[-c(5)])[[i]], las = 2, 
       tcl = -0.2, cex.axis = cxa, mgp = c(3, 0.5, 0))
  at <- seq(0, yls[2], 2)
  if(i == 1) axis(2, at, las = 2, tcl = -0.2, cex.axis = cxa, mgp = a2mgp)
  if(i == 2) {
    axis(2, at, labels = rep("", length(at)), las = 2, tcl = -0.2,
         cex.axis = cxa, mgp = a2mgp)
  }
  mtext(ylabs[i], side = 1, line = 3, cex = 0.9)
  if(i == 1) {
    mtext("Magnitude of difference between Extent/Duration", side = 2, 
          line = 2.2, cex = 0.9)
    mtext("and Actual Extent/Duration", side = 2, line = 1.4, cex = 0.9)
  }
  lrng <- c(min(di$xl) - 0.05 * min(di$xl), mean(di$xl))
  xs <- seq(lrng[1], lrng[2], by = (diff(lrng)) / (length(cols)))
  rect_shade(xs, y = c(13, 14), fillcol = cols, linecol = "grey30")
  text(xs, rep(12.5, length(xs)), di$wgti, srt = 90, col = "grey30", cex = cxa)
  text(mean(xs), 14.25, "% of observations", col = "grey30", cex = cxa)
  mtext(c("A", "B")[i], side = 3, line = 0, cex = 1, adj = 0.02)
}
o <- dev.off()


```

## Figure 3S

Box plots by bins of extent/duration *and* by observational type, colored and labelled by the percent of observations per bin.

```{r, eval = e2}
cxa <- 1
ayl <- c(0, 20)
byl <- c(0, 40)

xlabs <- c("Extent", "Duration")
ylabs <- c("All", "Field", "Automated", "Remote", "Paleo")
a2mgp <- c(3, 0.5, 0)
bwex <- c(0.9, 1.4)

png("figures/fig3s.png", width = 5, height = 10, res = 600, units = "in")
par(mfrow = c(5, 2), mar = c(0.5, 0.75, 0.5, 0.75), oma = c(5, 5, 0, 0))
for(i in 1:length(dif_ints)) {  # i <- 1
  for(j in 1:2) {  # j <- 1
    di <- dif_ints[[i]][[j]]  # subset list element
    bxa <- rowMeans(cbind(di$ival[-length(di$ival)], di$ival[-1]))[di$ints]
    bxw <- (di$ival[-1] - di$ival[-length(di$ival)])[di$ints]
    cols <- brewer.pal(pal_len[j], colpals[[j]])
    mus <- di$db[order(int), list("mu" = mean(dif)), by = int]  # means
    
    yls <- c(0, 14)
    plot(di$xl, yls, pch = "", xlab = "", ylab = "", xaxt = "n", yaxt = "n",
         mgp = c(3, 0.5, 0), tcl = -0.2, xaxs = "i")
    polygon(c(di$xl, rev(di$xl), di$xl[1]), c(-1, -1, rep(yls[2] * 1.2, 2), -1),             col = "grey90")
    ylines(yr = yls, by = 2, col = "grey95")
    b <- boxplot(formula = dif ~ int, data = di$db, col = cols[di$col], 
                 range = 0, width = bxw, at = bxa, xlim = di$xl, 
                 boxwex = bwex[j], axes = FALSE, outline = FALSE, lty = 1, 
                 add = TRUE, staplewex = 0.2)
    points(bxa, mus$mu, pch = 20, col = "black")
    text(bxa, b$stats[5, ] + 0.35, round(di$wgt, 1), cex = 0.8)
    
    xvals <- list(alab2, tlab2[-c(5)])[[j]]
    if(i == length(dif_ints)) {
      axis(1, at = ivals[[j]], labels = xvals, las = 2, tcl = -0.2, 
           cex.axis = cxa, mgp = c(3, 0.5, 0))
    } else {
      axis(1, at = ivals[[j]], labels = rep("", length(xvals)), las = 2,
           tcl = -0.2, cex.axis = cxa, mgp = c(3, 0.5, 0))
    }
    at <- seq(0, yls[2], 2)
    if(j == 1) {
      axis(2, at, las = 2, tcl = -0.2, cex.axis = cxa, mgp = a2mgp)
    } else {
      axis(2, at, labels = rep("", length(at)), las = 2, tcl = -0.2,
           cex.axis = cxa, mgp = a2mgp)
    }
    if(i == 5) mtext(xlabs[j], side = 1, line = 4, cex = cxa)
    if(j == 1) mtext(ylabs[i], side = 2, line = 4, cex = cxa)
    if(i == 3 & j == 1) {
      mtext(paste("Magnitude of difference between Extent/Duration",
                  "and Actual Extent/Duration"), side = 2, line = 1.4, 
            cex = 0.9)
    }
    if(i == 1) {
      lrng <- c(min(di$xl) - 0.1 * min(di$xl), mean(di$xl))
      xs <- seq(lrng[1], lrng[2], by = (diff(lrng)) / (length(cols)))
      rect_shade(xs, y = c(12.5, 13.5), fillcol = cols, linecol = "grey30")
      text(xs, rep(11.8, length(xs)), di$wgti, srt = 90, col = "grey30",
           cex = cxa)
      text(mean(xs), 14.1, "% of observations", col = "grey30", cex = cxa)
      
    }
  }
}
o <- dev.off()

```

## Observational type by bin

To see how much different observational types contribute to each bin.

```{r, eval = e}
dif_ints$all$e$db[order(int), .N, by = list(int, type)][, {
  ggplot() + aes(x = int, y = N, color = type) + geom_line() +
    scale_color_brewer(type = "div") +
    xlab("Extent interval")
}]

dif_ints$all$d$db[order(int), .N, by = list(int, type)][, {
  ggplot() + aes(x = int, y = N, color = type) + geom_line() + 
    scale_color_brewer(type = "div") +
    xlab("Duration interval")
}]


```

