---
title: "Main Analyses - Figure 2"
output: 
  rmarkdown::html_vignette:
    toc: yes
    number_sections: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{5. Main Analyses - Figure 2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', fig.height = 4, 
                      fig.width = 5, cache = TRUE)
```
```{r, echo=FALSE}
e <- TRUE
e2 <- FALSE
```


# Overview

The resampled full dataset created in the [Data Pre-processing vignette](data-pre-process.html) is used to construct kernel density estimates showing varying 2-D contrasts of observational scale dimensions.

# Analyses
## Data and libraries

```{r, message = FALSE, warning=FALSE}
library(ecoscales)
library(doMC)
library(viridis)
data("dimbreaks")
data("bootperturb")
```


<a href="#top">Back to top</a>

## Kernel density estimates

Kernel density estimates are created from juxtapositions of different scale dimensions, which are further broken down by observational types for supplementary analyses. 

The processing time is slow so this chunk is run separately, and results are added to package for lazy load, so that package vignettes can build more quickly (and to avoid hassles with caching chunks and vignettes). 

```{r, eval = FALSE}
bootp_dt <- rbindlist(bootperturb)

# choose which dataset to use
kdat <- copy(bootp_dt)  # bootstrap

# limits and variables
lims <- list("t_btwn_samp" = c(-5, 5.57), "plot_res" = c(-5, 8), 
             "eff_ext" = range(aaxis2$logarea),
             "eff_dur" = c(-5, 5.56))  # limits
xyvars <- list(c(x = "plot_res", y = "t_btwn_samp"),  # variable pairs 
               c(x = "eff_dur", y = "eff_ext"), 
               c(x = "plot_res", y = "eff_ext"), 
               c(x = "eff_dur", y = "t_btwn_samp"))
dropones <- c(TRUE, FALSE, FALSE, TRUE)

# Prepare datasets and run kernel density estimates
types <- c("field|remote|paleo|automated|other", "field", "automated", 
           "remote|other", "paleo")
# qtl <- function(x) quantile(x, probs = seq(0, 1, 0.1), na.rm = TRUE, type = 1)

registerDoMC(8)
kderla <- foreach(i = 1:length(types)) %:%   # i <- 1
  foreach(j = 1:length(xyvars)) %dopar% {  # j <- 1
    dfd <- kdat[like(type, types[i])]
    xyv <- xyvars[[j]]
    xl <- lims[[xyv["x"]]]
    yl <- lims[[xyv["y"]]]
    
    # create input spdf
    odt <- kdat_setup(dfd, x = xyv["x"], y = xyv["y"], xlim = xl, ylim = yl, 
                      drop.oneoff = dropones[j])

    # kernel densities
    kder <- kdensity(xl[1], xl[2], yl[1], yl[2], 0.1, odt$dat, 1)
    kder <- (kder / cellStats(kder, sum)) * 100
    list("kde" = kder, "mean" = odt$mu, "med" =  odt$med)
  }
kderla <- lapply(kderla, function(x) {
  names(x) <- c("resint", "durext", "resext", "durint")
  x
})
names(kderla) <- c("all", "field", "automated", "remote", "paleo")

save(kderla, file = "data/kderla.rda")
```

```{r, echo=FALSE}
data("kderla")
```

## Masks for impossible domains

To block out areas in subsequent plot that are not physically possible (e.g. interval > duration; resolution > extent)
```{r, eval = e}
# create masks for impossible values in res vs ext, and dur vs int
yr <- xr <- kderla$all$resext$kde
v <- xFromCell(xr, 1:ncell(xr))
vx <- log10(10^v / 10000)  # convert plot_res to ha 
vy <- yFromCell(yr, 1:ncell(yr))
xrng <- range(v[vx > vy])  # select values from v, not vx, because of scale chng
yrng <- range(vy[vy < vx])
pcrd <- cbind(c(xrng[1], xrng[2], xrng[2], xrng[1]), 
              c(yrng[1], yrng[1], yrng[2], yrng[1]))
msk3sp <- SpatialPolygons(list(Polygons(list(Polygon(coords = pcrd)), ID = 1)))
# plot(kderla$all$resext$kde)
# plot(msk3sp, add = TRUE, col = "blue", border = FALSE)

yr <- xr <- kderla$all$durint$kde
vx <- xFromCell(xr, 1:ncell(xr))
vy <- yFromCell(yr, 1:ncell(yr))
xrng <- range(vx[vx < vy])  # select values from vx
yrng <- range(vy[vy > vx])
pcrd <- cbind(c(xrng[1], xrng[2], xrng[1], xrng[1]), 
              c(yrng[2], yrng[2], yrng[1], yrng[2]))
msk4sp <- SpatialPolygons(list(Polygons(list(Polygon(coords = pcrd)), ID = 1)))
# plot(kderla$all$durint$kde)
# plot(msk4sp, add= TRUE)
```

<a href="#top">Back to top</a>

## Figure 2

Main figure of kernel densities, with median values added to density plots. 

```{r, eval=e2}
# global function for defining breaks
brkfun <- function(ext, ival, n) {
  rng <- range(ext[is.finite(ext)])
  bwidth <- (rng[2] - rng[1]) / n
  brks <- seq(rng[1], rng[2], bwidth)
  brklabs <- seq(0, round(rng[2], 2), ival)
  list("brks" = brks, "labs" = brklabs)
}

# global function for plotting legend
legfun <- function(kde, ival = 0.01, n, mgp = c(3, 0.25, 0), tcl = -0.1, 
                   cxa = cxa, ls = 0.9, col = cols, lw = 1.25, 
                   smallplot = c(0.875, 0.9, 0.4, 0.95)) {
  brks <- brkfun(ext = kde, ival = ival, n = n)  # breaks
  aargs <- list(mgp = mgp, at = brks$labs, labels = brks$labs, 
                cex.axis = cxa, tcl = tcl)
  plot(kde, legend.only = TRUE, axis.args = aargs, legend.width = lw,
       legend.shrink = ls, col = col, smallplot = smallplot)#inferno(cuts))
}
  
# cxa = 0.7
ll <- 3.5
cxa = 0.8
# stps <- sapply(c(3, 4), function(x) which(kdat$st == x))  # IDs rs/auto/paleo
# pchs <- list(1, "+")
cexs <- c(0.4, 0.6)
cuts <- 30
bump <- 3
cols <- inferno(cuts + bump)[-c(2:(bump + 1))]
# plot(1:cuts, col = cols, pch = 20)

axv <- list("A" = list("x" = aaxis1$logres[-1], "y" = taxis1$logdays[-c(5)]), 
            "B" = list("x" = taxis2$logdays[-5], "y" = taxis2$logdays[-5]),
            "C" = list("x" = aaxis1$logres[-1], "y" = aaxis2$logarea), 
            "D" = list("x" = taxis2$logdays[-5], "y" = aaxis2$logarea))
axl <- list("A" = list("x" = alab1[-1], "y" = tlab1[-c(5)]), 
            "B" = list("x" = tlab2[-5], "y" = tlab2[-5]), 
            "C" = list("x" = alab1[-1], "y" = alab2), 
            "D" = list("x" = tlab2[-5], "y" = alab2))
d <- c("Resolution", "Extent", "Interval", "Duration")
labs <- list("A" = d[c(1, 3)], "B" = d[4:3], "C" = d[1:2], "D" = d[c(4, 2)])
ptnms <- names(kderla)[-1]   # names for point plots
ppch <- c("f", "a", "r", "p")

png("figures/fig2.png", width = 7, height = 6, res = 600, units ="in")
par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
    mgp = c(2, 0.25, 0))
# resolutions
ki <- c(1, 4, 3, 2)  # index to account for change in kde order
for(i in 1:4) {  # i <- 1
  # plot main images and masks
  image(kderla$all[[ki[i]]]$kde, axes = FALSE, xlab = "", ylab = "", col = cols)
  if(i == 2) plot(msk4sp, add = TRUE, col = "grey30", border = "transparent")
  if(i == 3) plot(msk3sp, add = TRUE, col = "grey30", border = "transparent")
  axis(1, axv[[i]]$x, labels = axl[[i]]$x, las = 2, tcl = -0.2, cex.axis = cxa)
  axis(2, axv[[i]]$y, labels = axl[[i]]$y, las = 2, tcl = -0.2, cex.axis = cxa)
  
  # observation type means
  for(k in 1:length(ptnms)) {  # k <- 1
    # ppts <- kderla[[ptnms[k]]][[ki[i]]]$mean
    ppts <- kderla[[ptnms[k]]][[ki[i]]]$med
    if(i %in% c(2, 4)) ppts[x > axv[[i]]$x[9], x := axv[[i]]$x[9]]
    points(ppts, pch = ppch[k], col = "grey80", font = 2)
    # points(ppts2, pch = ppch[k], col = "green4")
  }    

  # legend and labels
  legfun(kderla$all[[ki[i]]]$kde, n = cuts, cxa = cxa, col = cols)
  for(j in 1:2) mtext(side = j, labs[[i]][j], cex = cxa, line = ll)
  mtext(LETTERS[i], side = 3, line = -0.1, cex = 0.8, adj = 0.025)
}
o <- dev.off()

```

## Figure 2S

Expanded version of Figure 2, showing kernel densities of different observational categories. 

```{r, eval = e2}
ylabs <- c("All", "Field", "Automated", "Remote", "Paleo")
titles <- c("Res v Int", "Dur v Int", "Res v Ext", "Dur v Ext")

png("figures/fig2s.png", width = 8.5, height = 9, res = 600, 
    units = "in")
par(mfrow = c(5, 4), mar = c(4.5, 4.5, 0.3, 2.5), oma = c(0, 1, 1, 0), 
    mgp = c(2, 0.25, 0))
kj <- c(1, 4, 3, 2)  # index to account for change in kde order
for(i in 1:5) {
  for(j in 1:4) {  # i <- 1
    # plot main images and masks
    image(kderla[[i]][[kj[j]]]$kde, axes = FALSE, xlab = "", ylab = "", 
          col = cols)
    if(j == 2) plot(msk4sp, add = TRUE, col = "grey30", border = "transparent")
    if(j == 3) plot(msk3sp, add = TRUE, col = "grey30", border = "transparent")
    axis(2, axv[[j]]$y, labels = axl[[j]]$y, las = 2, tcl = -0.2,
         cex.axis = cxa)
    axis(1, axv[[j]]$x, labels = axl[[j]]$x, las = 2, tcl = -0.2,
         cex.axis = cxa)
    if(j == 1) mtext(ylabs[i], side = 2, line = 4, cex = cxa)
    if(i == 1) mtext(titles[j], side = 3, line = 0, cex = cxa)

    # legend and labels
    legfun(kderla$all[[kj[j]]]$kde, n = cuts, cxa = cxa, col = cols, 
           smallplot = c(0.85, 0.875, 0.4, 0.95))
  }
}
o <- dev.off()

```




