---
title: "Main Analyses - Figure 1"
output: 
  rmarkdown::html_vignette:
    toc: yes
    number_sections: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Main Analyses - Figure 1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', fig.height = 4, 
                      fig.width = 5)
```

# Overview

The resampled full dataset created in the [Data Pre-processing vignette](data-pre-process.html) are used to create histograms of scale in each of the 4 assessed dimensions

# Analyses

## Data and libraries
```{r}
library(ecoscales)
data("datf")
data("dimbreaks")
data("bootperturb")
```


## Histograms

Constructed from `bootperturb`. 
```{r hists, cache=TRUE}
# break points for data
brksl <- list(aaxis1$logres, aaxis2$logarea, taxis1$logdays, taxis2$logdays)  

# Assign bootstrapped values to bins, making necessary adjustments as needed 
vs <- c("res", "ext", "int", "dur")
hdat_bs <- lapply(1:length(bootperturb), function(x) {  # x <- 1
  # print(x)
  DT <- copy(bootperturb[[x]])
  hdb <- DT[, list("type" = type, "res" = log10(plot_res), 
                   "ext" = log10(eff_ext), "int" = log10(t_btwn_samp), 
                   "dur" = log10(eff_dur))]
  hdb[res < vlr[1], res := vlr[1]]  # set min res to 0.01 cm"^2 
  hdb[res > vlr[2], res := vlr[2]]  # set max res to 10000 ha 
  hdb[ext < vle[1], ext := vle[1]]  # set minimum extent -6 (0.01 m2)
  hdb[ext > vle[2], ext := vle[2]]  # set maximum extent 10 (10^10 ha)

  # histograms across all types and by type  
  types <- c("field|automated|remote|other|paleo", "field", "automated",
             "remote|other", "paleo")
  histso <- lapply(1:length(types), function(k) {  # k <- 5
    dat <- hdb[like(type, types[k])]
    hists <- lapply(1:length(vs), function(y) {  # y <- 2
      #print(y)
      datv <- dat[[vs[y]]]
      hbrks <- brksl[[y]]
      datv[datv < hbrks[1]] <- hbrks[1]
      datv[datv > hbrks[length(hbrks)]] <- hbrks[length(hbrks)]
      h <- hist(datv, breaks = hbrks, plot = FALSE)
      # h$density <- h$counts / sum(h$counts) * 100
      h$density <- h$counts / 378 * 100  # calculate as % of total observations
      h
    })
    names(hists) <- vs
    hists
  })
  names(histso) <- c("all", "field", "automated", "remote", "paleo")
  histso
})

# Reduce dataset to just density values
typenms <- c("all", "field", "automated", "remote", "paleo")
hdat_bsdt <- lapply(typenms, function(j) { # j <- "field"
  hdat_bsi <- lapply(1:length(vs), function(k) { # k <- 1
    mat <- do.call(rbind, lapply(1:length(hdat_bs), function(i) {  # i <- 1
      hdat_bs[[i]][[j]][[k]]$density
    }))
    DT <- data.table(mat)
  })
  names(hdat_bsi) <- vs
  hdat_bsi
})
names(hdat_bsdt) <- typenms

# And calculate mean, 2.5th, 97.5th from observations
qtf <- function(x) c(mean(x), quantile(x, probs = c(0.025, 0.975))) # quant f
hdatstat_bt <- lapply(1:length(hdat_bsdt), function(x) {
  hdatstat_bti <- lapply(1:length(vs), function(y) {
    h <- hdat_bsdt[[x]][[y]][, lapply(.SD, qtf)]
    hmu <- unlist(h[1, ])
    h2 <- unlist(h[2, ])
    h98 <- unlist(h[3, ])
    list("mu" = hmu, "p2" = h2, "p98" = h98)
  })
  names(hdatstat_bti) <- vs
  hdatstat_bti
})
names(hdatstat_bt) <- typenms
```

<a href="#top">Back to top</a>


## Figure 1 (main text)
```{r, message=FALSE, warning=FALSE}
xlabs <- c("Resolution", "Extent", "Interval", "Duration")
ylabs <- c("All", "Field", "Automated", "Remote", "Paleo")
axes2 <- list("aax1" = aaxis1$logres, "aax2" = aaxis2$logarea, 
              "tax1" = c(taxis2$logdays[-c(10:11)],
                         mean(taxis2$logdays[c(10:11)])), 
              "tax2" = taxis2$logdays)
axlabs <- list(alab1, alab2, tlab1[-10], tlab2)
mga <- -0.34
sfigl <- -1
cxa <- 0.8
cxl <- 0.9
reds <- brewer.pal(9, name = "Reds")
blues <- brewer.pal(9, name = "Blues")
# plot(1:2, pch = 20, col = blues[c(4, 6)])
yl <- 35
cols <- c(reds[c(4, 6)], blues[c(4, 6)])

# pdf("paper/figures/hists_bs.pdf", width = 5, height = 5)
png("figures/fig1.png", width = 5, height = 5, res = 600, 
    units = "in")
par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
for(i in 1:length(vs)) {  # i <- 3
  hvals <- hdatstat_bt$all[[i]]
  axv <- brksl[[i]]
  mids <- sapply(2:length(axv), function(x) mean(c(axv[x], axv[x - 1])))
  
  axvw <- axv[-1] - axv[-length(axv)]  # bar widths
  if(i %in% c(1:2, 4)) {
    xaxp <- cumsum(c(0, axvw))  # x axis positions
  } else if(i == 3) {
    xaxp <- cumsum(c(0, axvw[-length(axvw)]))  # x axis positions
    xaxp[length(xaxp)] <- xaxp[length(xaxp)] + axvw[length(axvw)] / 2
  }

  # base plot
  yl <- ifelse(i %in% c(1, 2), 25, 40) 
  b <- barplot(hvals$mu, width = axvw, space = c(0, 0), col = cols[i], 
               yaxt = "n", ylim = c(0, yl), border = "grey40", 
               axisnames = FALSE, xlab = xlabs[i])
  axis(1, at = xaxp, labels = axlabs[[i]], las = 2, 
       cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.25, 0))
  axis(2, las = 2, cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.25, mga))
  if(i %in% c(1, 3)) {
    mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
  }
  mtext(LETTERS[i], side = 3, line = sfigl, cex = 0.8, adj = 0.05)

  # confidence intervals
  points(b, hvals$p2, pch = "-", cex = 1, col = "grey40")
  points(b, hvals$p98, pch = "-", cex = 1, col = "grey40")
}
dev.off()
```

<a href="#top">Back to top</a>


## Figure 1s

Detailed by observation type, for supplemental material.

```{r, message=FALSE, warning=FALSE}
xlabs <- c("Resolution", "Extent", "Interval", "Duration")
ylabs <- c("All", "Field", "Automated", "Remote", "Paleo")
axes2 <- list("aax1" = aaxis1$logres, "aax2" = aaxis2$logarea, 
              "tax1" = c(taxis2$logdays[-c(10:11)],
                         mean(taxis2$logdays[c(10:11)])), 
              "tax2" = taxis2$logdays)
axlabs <- list(alab1, alab2, tlab1[-10], tlab2)
mga <- -0.5
sfigl <- -1.2
cxa <- 1.3#0.8
cxl <- 1.3#0.9
reds <- brewer.pal(9, name = "Reds")
blues <- brewer.pal(9, name = "Blues")
# plot(1:2, pch = 20, col = blues[c(4, 6)])
yl <- c(40, 40, 6, 6, 6)
cols <- c(reds[c(4, 6)], blues[c(4, 6)])

# pdf("paper/figures/hists_bs.pdf", width = 5, height = 5)
png("figures/fig1s.png", width = 9, height = 11, res = 600, 
    units = "in")
par(mfrow = c(5, 4), mar = c(1, 0.25, 0.25, 0), oma = c(6, 5, 1, 1))
for(i in 1:length(hdatstat_bt)) {  # i <- 5
  for(j in 1:length(vs)) {  # j <- 1
    hvals <- hdatstat_bt[[i]][[j]]
    axv <- brksl[[j]]
    
    axvw <- axv[-1] - axv[-length(axv)]  # bar widths
    if(j %in% c(1:2, 4)) {
      xaxp <- cumsum(c(0, axvw))  # x axis positions
    } else if(j == 3) {
      xaxp <- cumsum(c(0, axvw[-length(axvw)]))  # x axis positions
      xaxp[length(xaxp)] <- xaxp[length(xaxp)] + axvw[length(axvw)] / 2
    }

    # base plot
    b <- barplot(hvals$mu, width = axvw, space = c(0, 0), col = cols[j], 
                 yaxt = "n", ylim = c(0, yl[i]), border = "grey40", 
                 axisnames = FALSE)
    
    # Axes and labels
    if(i == 5) {  # main x axis with values (at bottom)
      axis(1, at = xaxp, labels = axlabs[[j]], las = 2, 
           cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.5, 0))
    } 
    if(i %in% 1:4) {  # ticks for other x axes
      axis(1, at = xaxp, labels = rep("", length(axes2[[j]])), las = 2, 
           cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.5, 0))
    }
    if(j == 1) {  # y axes, including label for 3rd row
      axis(2, las = 2, cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.5, mga))
      mtext(ylabs[i], side = 2, line = 3.25, cex = cxa)
    }
    
    # differentiate scale change for rows 3-5 with color
    smaxcol <- brewer.pal(9, "Oranges")[6]  # show scale for 3-5 on 1-2
    if(j == 1) {  
      axis(2, at = c(0, 6), labels = c("", 6), las = 2, col.ticks = smaxcol,
           col.axis = smaxcol, col = smaxcol, cex.axis = cxa, tcl = -0.2, 
           mgp = c(2, 0.5, mga))
    }
    if(j == 1 & i %in% 3:5) {  # axes for 3:5
      axis(2, las = 2, cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.5, mga), 
           col.ticks = smaxcol, col.axis = smaxcol, col = smaxcol)
    }
    if(i == 1) mtext(xlabs[j], side = 3, line = -0.25, cex = cxa)
    if(i == 3 & j == 1) {
      mtext(expression(italic("Percent of observations")), side = 2, line = 1.5,
            cex = 1)
    }

    # add 95% confidence intervals
    points(b, hvals$p2, pch = "-", cex = 1.5, col = "grey40")
    points(b, hvals$p98, pch = "-", cex = 1.5, col = "grey40")
    
  }
}
dev.off()
```

<a href="#top">Back to top</a>


```{r, echo = FALSE, eval = FALSE}
# save(hdat, hdatstat_bt, file = "data/hdatstat.rda")
```


## Figure 1 summary statistics 

Statistics associated with Figure 1: percentage of observations falling within different scale ranges, and associated with different observational methods

```{r}
# res
hdatst <- hdatstat_bt$all$res$mu
l <- length(hdatst)
hds <- c("10cm^2 - 1 m^2", "<1m^2", "1 m^2-1 ha", "1-10000 ha")
inds <- list(4:6, 1:6, 7:10, 11:l)
sapply(1:length(inds), function(x) {
  v <- round(sum(hdatst[inds[[x]]]), 1)
  names(v) <- hds[x]
  v
})

# extent
# hdatst <- hdatstat[[2]]$density
hdatst <- hdatstat_bt$all$ext$mu
l <- length(hdatst)
hds <- c("<1 ha", "<10 ha", "<100 ha", "<1000ha", "<10Kha", "10-1k ha", 
         "1k-10k", "10k-100k", "100k-1Mk", ">1Kha", ">100Kha", ">1Mha")
inds <- list(1:6, 1:7, 1:8, 1:9, 1:10, 8:9, 10, 11, 12, 10:l, 12:l, 13:l)
sapply(1:length(inds), function(x) {
  v <- round(sum(hdatst[inds[[x]]]), 1)
  names(v) <- hds[x]
  v
})

# interval
# hdatst <- hdatstat[[3]]$density
hdatst <- hdatstat_bt$all$int$mu
l <- length(hdatst)
hds <- c("norep", "<daily", "<weekly", "<minute", "min-hour", 
         "hr-day", "day-wk", "wk-mo", "mo-yr", "yr-dec", "dec-cen", "cen-mil")
inds <- list(l, 1:3, 1:4, 1, 2, 3, 4, 5, 6, 7, 8, 9)
sapply(1:length(inds), function(x) {
  v <- round(sum(hdatst[inds[[x]]]), 1)
  names(v) <- hds[x]
  v
})

# duration
# hdatst <- hdatstat[[4]]$density
hdatst <- hdatstat_bt$all$dur$mu
l <- length(hdatst)
hds <- c("<=daily", "day-mo", "mo-yr", "yr-dec", "dec-cent")
inds <- list(1:3, 4:5, 6, 7, 8:10)
sapply(1:length(inds), function(x) {
  v <- round(sum(hdatst[inds[[x]]]), 1)
  names(v) <- hds[x]
  v
})
```

<a href="#top">Back to top</a>

```{r, echo = FALSE, eval = FALSE}
# data checks (run in chunk only)
# recreate plots from datf to compare visually to plots created in fig 1s, to
# make sure they look more or less the same
axlabs <- list(alab1, alab2, tlab1[-10], tlab2)
dfnms <- c("plot_res", "eff_ext", "t_btwn_samp", "eff_dur")
reds <- brewer.pal(9, name = "Reds")
blues <- brewer.pal(9, name = "Blues")
cols <- c(reds[c(4, 6)], blues[c(4, 6)])
cxa <- 0.8
cxl <- 0.9

par(mfrow = c(2, 2))
pat <- "paleo"
yl <- 6
for(i in 1:length(dfnms)) {  # i <- 1
  print(dfnms[i])
  if(i %in% c(1:2, 4)) {
    datv <- datf[like(study_type, pat), log10(get(dfnms[i]))]    
  }
  if(i == 3) {
    datv <- datf[like(study_type, pat),
                 log10(to_infin_byond(get(dfnms[i])))]    
  }
  hbrks <- brksl[[i]]
  datv[datv < hbrks[1]] <- hbrks[1]
  datv[datv > hbrks[length(hbrks)]] <- hbrks[length(hbrks)]
  h <- hist(datv, breaks = hbrks, plot = FALSE)
  # h$density <- h$counts / sum(h$counts) * 100
  h$density <- h$counts / 378 * 100  # calculate as % of total observations

  axv <- brksl[[i]]
  axvw <- axv[-1] - axv[-length(axv)]  # bar widths
  axvw <- axv[-1] - axv[-length(axv)]  # bar widths
  if(i %in% c(1:2, 4)) {
    xaxp <- cumsum(c(0, axvw))  # x axis positions
  } else if(i == 3) {
    xaxp <- cumsum(c(0, axvw[-length(axvw)]))  # x axis positions
    xaxp[length(xaxp)] <- xaxp[length(xaxp)] + axvw[length(axvw)] / 2
  }
  b <- barplot(h$density, width = axvw, space = c(0, 0), col = cols[i], 
               yaxt = "n", ylim = c(0, yl), border = "grey40", 
               axisnames = FALSE)
  axis(1, at = xaxp, labels = axlabs[[i]], las = 2, 
       cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.25, 0))
  axis(2, las = 2, cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.25, 0))
}

# Data check - are we extracting the statistics correctly? Check on subsets
l <- list(c("field", "field", 2), c("paleo", "paleo", 4), 
          c("remote|other", "remote", 3), c("automated", "automated", 1))
for(j in 1:length(l)) {
  types <- l[[j]][1]
  types2 <- l[[j]][2]
  i <- as.numeric(l[[j]][3])
  vss <- vs[i]
  tst <- lapply(1:length(bootperturb), function(x) {  # x <- 1
    DT <- copy(bootperturb[[x]])
    hdb <- DT[, list("type" = type, "res" = log10(plot_res),
                     "ext" = log10(eff_ext), "int" = log10(t_btwn_samp),
                     "dur" = log10(eff_dur))]
    hdb[res < vlr[1], res := vlr[1]]  # set min res to 0.01 cm"^2 
    hdb[res > vlr[2], res := vlr[2]]  # set max res to 10000 ha 
    hdb[ext < vle[1], ext := vle[1]]  # set minimum extent -6 (0.01 m2)
    hdb[ext > vle[2], ext := vle[2]]  # set maximum extent 10 (10^10 ha)
    dat <- hdb[like(type, types)]
    datv <- dat[[vss]]
    hbrks <- brksl[[i]]
    datv[datv < hbrks[1]] <- hbrks[1]
    datv[datv > hbrks[length(hbrks)]] <- hbrks[length(hbrks)]
    h <- hist(datv, breaks = hbrks, plot = FALSE)
    h$density <- h$counts / 378 * 100  # calculate as % of total observations
    h$density
  })
  print(sapply(sample(1:1000, size = 20), function(x) {
    identical(hdat_bs[[x]][[types2]][[i]]$density, tst[[x]])
  }))
  
  # ran with permutation above--all check out
  q2 <- function(x) quantile(x, 0.025)
  q98 <- function(x) quantile(x, 0.975)
  a <- identical(unname(hdatstat_bt[[types2]][[i]]$mu), 
                 colMeans(do.call(rbind, tst)))
  b <- identical(unname(hdatstat_bt[[types2]][[i]]$p2), 
                 apply(do.call(rbind, tst), 2, q2))
  d <- identical(unname(hdatstat_bt[[types2]][[i]]$p98), 
                 apply(do.call(rbind, tst), 2, q98))
  print(c(a, b, d))
}

# all checks true
```

```{r, echo = FALSE}
rm(list = ls())
```
