---
title: "Sensitivity-related Analyses"
output: 
  rmarkdown::html_vignette:
    toc: yes
    number_sections: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{8. Sensitivity Analyses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  # collapse = TRUE,
  comment = "#>"
)

```
```{r, echo = FALSE}
e <- TRUE
e2 <- TRUE
```

# Overview

These analyses examine two further specific aspects of uncertainty in the study's methods: 

1. The percent of studies in which the scale of particular dimensions had to be estimated;
2. The impact of bandwidth on kernel density estimates 

# Analyses
## Data and libraries
```{r, message=FALSE, warning=FALSE}
library(ecoscales)
library(viridis)
library(doMC)
data("bootperturb")
data("dimbreaks")
data("datf")

f <- system.file("extdata/main-results.xlsx", package = "ecoscales")
dat <- data.table(read_excel(f))
```

## Proportion of observations estimated

This assessment is based on notes that each observer made in their data sheets specifying which dimensions for each observation were uncertain (and thus had to be estimated) based on the description of methods in the paper in question.  These notes occur in "sensitivity" column of the main data sheet. 

The notes themselves are noisy and need to be wrangled to get them into the correct dimensional names. 

```{r, eval = e, message=FALSE}

svars <- c("plot_res", "n_sites", "act_ext", "eff_ext", "samp_dur",
           "t_btwn_samp", "act_dur", "eff_dur")
sens <- gsub("\\\\", ";", dat$sensitivity)
sens <- gsub(", ", ";", gsub("; ", ";", sens))
sens <- tolower(sens)
sensl <- strsplit(sens, ";")
# unique(unlist(sensl))

sens2 <- gsub("plo_|plot-", "plot_", sens)  # fix this one
sens2 <- gsub("[[:blank:]]", "_", sens2)  # replace spaces with _
sens2 <- gsub("-", "_", sens2)  # replace - with _

# correction table--unique strings in sensitivity column and their replacements
corrtab <- cbind(c("^res", "plo_res", "plot_resolution", "plot_size", 
                   "sample_area", "samples_area", "sampled area", "study_area",
                   "sampled_area",
                   "t_btwn_samples", "t_btwn_sample", "t_btwn_samp_samp", 
                   "t_twn_samp", "t_btw_samp", "time_between_sampling", 
                   "time_between_samples", 
                   "samp_duratiion", "sampling_duration", "sample_duration", 
                   "number_of_sites", 
                   "study_span", "sutdy_span", "study_dur", "study_soan",
                   "samp_study", ":", " ", ";_", 
                   "duration", "dure", 
                   "extent"), 
                 c(rep("plot_res", 4), rep("act_ext", 5), rep("t_btwn_samp", 7),                    rep("samp_dur", 3), "n_sites", rep("eff_dur", 4), 
                   "samp;study", ";", ";", ";", 
                   "dur", "dur", "ext"))

# replace
for(i in 1:nrow(corrtab)) {
  sens2 <- gsub(corrtab[i, 1], corrtab[i, 2], sens2)
}

sens2[is.na(sens2)] <- "none"
sensl2 <- lapply(1:length(sens2), function(x) strsplit(sens2[x], ";")[[1]])

# create a sensitivity table, based on whether each observation has one of the 
# variables listed as uncertain
senst <- do.call(cbind.data.frame, lapply(svars, function(x) {  # x <- svars[1]
  v <- sapply(sensl2, function(y) ifelse(any(y == x), 1, 0)) # y <- sensl2[[37]]
  # ifelse(is.na(v), 0, v)
}))
colnames(senst) <- c("res", "n", "aext", "eext", "sampd", "tbtwn", "actdur", 
                     "effdur")

# apply fixes (mostly to account for observer omissions)
# 1. if samp_dur OR t_btwn_samp is uncertain, act_dur must also be. 
id <- which((senst$sampd == 1 | senst$tbtwn == 1) & senst$actdur == 0)
senst[id, "actdur"] <- 1

# 2. if act_dur is uncertain but t_btwn and samp_dur are not, treat
# both as uncertain
# although it is possible that just n_repeats is uncertain, but more likely 
# that observers didn't note this is sensitive. 
id <- which(senst$tbtwn == 0 & senst$sampd == 0 & senst$actdur == 1)
senst[id, c("sampd", "tbtwn")] <- 1

# 3. if plot_res or n_sites is uncertain, then act_ext must be
id <- which((senst$res == 1 | senst$n == 1) & senst$aext == 0)
senst[id, "aext"] <- 1

# percentage of records that are uncertain
knitr::kable(t(round(colSums(senst) / nrow(senst) * 100, 1)))
```

## Observations Assigned Token 1 day

Calculating the proportions of observations assigned this value overall, focusing on the `samp_dur` parameter
```{r}
v <- round(datf[samp_dur == 1, .N] / datf[, .N] * 100, 1)
v2 <- round(datf[samp_dur == 1 & study_type != "automated" & t_btwn_samp != 0, 
                 .N] / datf[, .N] * 100, 1)
knitr::kable(cbind(v, v2), col.names = c("% of total (all 1 day values)", 
                                         "% of total (no automated or one-off"))
```

<a href="#top">Back to top</a>

## Kernel Bandwidth

Testing the results on images related to varying kernel sizes.

### Run kernel density estimates

The processing time is slow so this chunk is run separately, and results are added to package for lazy load, so that package vignettes can build more quickly (and to avoid hassles with caching chunks and vignettes).  

```{r, eval = FALSE}
bootp_dt <- rbindlist(bootperturb)
kdat <- copy(bootp_dt)  # bootstrap

# limits and variables
lims <- list("t_btwn_samp" = c(-5, 5.57), "plot_res" = c(-5, 8), 
             "eff_ext" = range(aaxis2$logarea),
             "eff_dur" = c(-5, 5.56))  # limits
xyvars <- list(c(x = "plot_res", y = "t_btwn_samp"),  # variable pairs 
               c(x = "eff_dur", y = "eff_ext"))
dropones <- c(TRUE, FALSE, FALSE, TRUE)

# Prepare datasets and run kernel density estimates
registerDoMC(8)
kderls <- foreach(i = 1:length(xyvars)) %:%   # i <- 1
  foreach(j = c(0.4, 0.7, 1)) %dopar% {  # j <- 1
    xyv <- xyvars[[i]]
    xl <- lims[[xyv["x"]]]
    yl <- lims[[xyv["y"]]]
    
    # create input spdf
    odt <- kdat_setup(kdat, x = xyv["x"], y = xyv["y"], xlim = xl, ylim = yl, 
                    drop.oneoff = dropones[i])

    # kernel densities
    kder <- kdensity(xl[1], xl[2], yl[1], yl[2], 0.1, odt$dat, j)
    kder <- (kder / cellStats(kder, sum)) * 100
    kder
  }
kderls <- lapply(kderls, function(x) {
  names(x) <- paste0("k", c(0.4, 0.7, 1))
  x
})
names(kderls) <- c("resint", "durext")

save(kderls, file = "data/kderls.rda")
```

<a href="#top">Back to top</a>

### Plot
```{r, echo = FALSE}
data("kderls")
```
```{r, eval = e2}
# global function for defining breaks and plotting legend (haven't added to 
# package since used <3 times!)
# breaks
brkfun <- function(ext, ival, n) {
  rng <- range(ext[is.finite(ext)])
  bwidth <- (rng[2] - rng[1]) / n
  brks <- seq(rng[1], rng[2], bwidth)
  brklabs <- seq(0, round(rng[2], 2), ival)
  list("brks" = brks, "labs" = brklabs)
}

# legend
legfun <- function(kde, ival = 0.01, n, mgp = c(3, 0.25, 0), tcl = -0.1, 
                   cxa = cxa, ls = 0.9, col = cols, lw = 1.25, 
                   smallplot = c(0.875, 0.9, 0.1, 0.8)) {
  brks <- brkfun(ext = kde, ival = ival, n = n)  # breaks
  aargs <- list(mgp = mgp, at = brks$labs, labels = brks$labs, 
                cex.axis = cxa, tcl = tcl)
  plot(kde, legend.only = TRUE, axis.args = aargs, legend.width = lw,
       legend.shrink = ls, col = col, smallplot = smallplot)#inferno(cuts))
}

# plot parameters
cxa <- 1.25
cuts <- 40
bump <- 3
cols <- inferno(cuts + bump)[-c(2:(bump + 1))]

axv <- list("A" = list("x" = aaxis1$logres[-1], "y" = taxis1$logdays[-c(5)]), 
            "B" = list("x" = taxis2$logdays[-5], "y" = aaxis2$logarea))
axl <- list("A" = list("x" = alab1[-1], "y" = tlab1[-c(5)]), 
            "D" = list("x" = tlab2[-5], "y" = alab2))

# As separate plots
# pdf("vignettes/figures/res_v_extent_ksize2.pdf", width = 8, height = 10)
png("figures/figS6.png", width = 8, height = 10, 
    units = "in", res = 600)
par(mfrow = c(3, 2), oma = c(6, 3, 0, 0), mar = c(1, 5, 3, 4))
for(i in 1:3) {  # i <- 1
  for(j in 1:2) { # j <- 1
    image(kderls[[j]][[i]], col = cols, axes = FALSE, xlab = "", ylab = "")
    if(i == 3) {
      axis(1, axv[[j]]$x, labels = axl[[j]]$x, las = 2, tcl = -0.2,
           cex.axis = cxa)
    }
    axis(2, axv[[j]]$y, labels = axl[[j]]$y, las = 2, tcl = -0.2, cex.axis=cxa)
    legfun(kderls[[j]][[i]], n = cuts, cxa = cxa, col = cols)
  }
}
o <- dev.off()

```

<a href="#top">Back to top</a>




